LME-MPM applied to quasi-brittle fracture.

### 2020-06-30 2020_EFM_MPM_Eigensoftening.tex
uGIMP

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025 $[L]$. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

\MMP{Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows that the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is released dinamically by the crack, the waves propagate
  along the domain.}

@@

\DM{The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle
materials and the stress oscillations observed with conventional
\acrshort{mpm}/\acrshort{gimp} are faced.} \PNA{Based on the
EigenMPM approach to fracture, two additional improvements have been proposed in the present research in order to solve the shortcomings of the original concept pointed out in the aforementioned paper \cite{Zhang_EE_2020}.}

Firstly, by introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrshort{fem}. Furthermore, this failure
criterion admits the design of more elaborated damage curves, which
can be employed in order to describe the failure process in more
sophisticated materials. Secondly to achieve better results, three new
numerical algorithms has been implemented. The \acrshort{lme}
approximants are employed to remove from
the solution those wiggles provoked by inaccuracies in the stress
integration. The traditional time scheme of the
\acrshort{mpm} has been improved with a Newmark Predictor Corrector
scheme as well. And the traditional concept of element connectivity
has been removed to allow a more flexible and efficient implementation of search
algorithm. In this research, the aforementioned techniques have helped
to improve the results in fracture propagation through the
eigensoftening and eigenerosion algorithms.

>>>>

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation and \acrshort{ugimp}. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current
simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025 $[L]$. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

\MMP{Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows that the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than \acrshort{lme}. This can be attributed to imprecision in the
stress field due to grid-crossing phenomena. \MMP{Alternative to
  \acrshort{lme}, \acrshort{ugimp} shape
functions can be employed to mitigate grid-crossing instabilities.
Figure \ref{fig:Reactions-cracked-panel-mode-I} support this since it
does not exhibits it when particles cross element borders as occur
with linear interpolation. Unfortunately the extended support of this
kind of shape function interfere with nodes in both sides of the notch,
producing an undesired ``tie-effect''. This involves in a violation of
the free border boundary condition in the notch. Possible solutions to this effect
are the suppression of the elements in the notch, within a CRAMP-like
method where two pairs of nodes are employed, or reducing the mesh size. With \acrshort{lme}
approximants these artefacts are not required since it is possible to control the support
of the shape function reducing the grid-crossing and avoiding the
``tie-effect'' in the notch at the same time.}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
remark clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the
  differences in the obtained peaks are not extremely significant, it
  can be owing to the aforementioned hookean material
  employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
  models will be employed, severe inaccuracies may appear in the
  stress field that could affect dramatically to the final 
  result. Focusing on the post failure behavior, it can be seen how
  \acrshort{lme} produces soft stress field evolution even after
  breaking.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the
  energy is released dinamically by the crack, the waves propagate
  along the domain.}

@@

\DM{The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle
materials and the stress oscillations observed with conventional
\acrshort{mpm} are faced.} \PNA{Based on the
EigenMPM approach to fracture, two additional improvements have been proposed in the present research in order to solve the shortcomings of the original concept pointed out in the aforementioned paper \cite{Zhang_EE_2020}.}

Firstly, by introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrshort{fem}. Furthermore, this failure
criterion admits the design of more elaborated damage curves, which
can be employed in order to describe the failure process in more
sophisticated materials. Secondly to achieve better results, three new
numerical algorithms has been implemented. The \acrshort{lme}
approximants are employed to remove from
the solution those wiggles provoked by inaccuracies in the stress
integration and \MMP{to avoid the ``tie-effect'' observed in the
\acrshort{ugimp}}. The traditional time scheme of the \acrshort{mpm}
has been improved with a Newmark Predictor Corrector scheme as
well. And the traditional concept of element connectivity has been
removed to allow a more flexible and efficient implementation of
search algorithm. In this research, the aforementioned techniques have helped
to improve the results in fracture propagation through the
eigensoftening and eigenerosion algorithms.
